% ============================================================================
% Appendix A1 — Positive Invariance of the State Space
% ============================================================================
\section{Positive Invariance of the State Space}
\label{app:invariance}

We verify that the physically meaningful region
$[0,\Vmax]\times[0,\infty)$ is forward invariant under
\eqref{eq:dVdt}--\eqref{eq:complexity}, and record the boundary facts used
throughout the paper. The right-hand side is locally Lipschitz on the
region (the barrier $V/(V+\varepsilon)$ is smooth for $V>-\varepsilon$), so
solutions exist and are unique; invariance follows from the sign of the
vector field on each boundary face.

\paragraph{Face $V=0$ (vitality floor).}
At $V=0$ the barrier factor vanishes, $V/(V+\varepsilon)=0$, so the drain
term of \eqref{eq:dVdt} is identically zero and
\[
\dot V\big|_{V=0} \;=\; R \;>\; 0
\qquad\text{for every } C\ge0 .
\]
The vector field points strictly \emph{into} the region: $V=0$ is repelling.
This is the fact that falsifies the ``finite-time depletion to zero''
theorem of earlier drafts and motivates its replacement,
Proposition~\ref{prop:band}. The companion test
\texttt{test\_dv\_dt\_positive\_at\_V\_zero} pins
$\dot V|_{V=0}=R$ exactly, for several values of $C$.

\paragraph{Face $V=\Vmax$ (vitality ceiling).}
At $V=\Vmax$ the recovery term vanishes and, since $O\ge O_0>0$,
\[
\dot V\big|_{V=\Vmax}
\;=\; -\,\delta\,O^{\gamma}\,\frac{\Vmax}{\Vmax+\varepsilon}
\;<\;0 ,
\]
so trajectories cannot cross the ceiling. (Numerically, integration from
$V_0=\Vmax$ respects this to solver tolerance;
\texttt{test\_simulate\_vitality\_stays\_below\_ceiling}.)

\paragraph{Face $C=0$ (capital axis).}
By \eqref{eq:impact}--\eqref{eq:dCdt}, $\dot C = C\bigl(\alpha
V/(1+\varphi O)-\mu\bigr)$, hence
\[
\dot C\big|_{C=0} \;=\; 0
\qquad\text{for every } V ,
\]
so $\{C=0\}$ is an \emph{invariant set}, not merely a reflecting boundary:
capital can neither become negative nor be created from nothing. Dynamics
restricted to the axis reduce to the scalar vitality equation at
$O=O_0$, whose unique rest point $V\approx9.934$ (the positive root of
\eqref{eq:vqe-quadratic} at $O=O_0$) is analyzed in
Appendix~\ref{app:regularization}, where it is shown to be a saddle of the
planar system.

\paragraph{Consequences.}
(i)~$[0,\Vmax]\times[0,\infty)$ is forward invariant; no clamping is needed
in exact arithmetic (the numerical clamps in the companion code are
solver-safety devices only).
(ii)~Because $\{C=0\}$ is invariant, no interior trajectory can reach it in
finite time, and any global-stability statement must be made on the open
set $\{C>0\}$---as in Theorem~\ref{thm:gas}.
(iii)~The barrier constant $\varepsilon$ trades regularity for a small
displacement of equilibrium quantities; the displacement is quantified in
Appendix~\ref{app:epsilon}.
