% ============================================================================
% Appendix A10 — Global Asymptotic Stability: Full Proof
% ============================================================================
\section{Global Asymptotic Stability: Full Proof of Theorem~\ref{thm:gas}}
\label{app:gas}

We assemble the complete proof from four ingredients: (1)~a forward-invariant
trapping rectangle with the \emph{corrected} ceiling $\Ctrap$;
(2)~a Bendixson--Dulac certificate excluding closed orbits;
(3)~uniqueness and local stability of the interior equilibrium; and
(4)~the boundary structure of the invariant axis $\{C=0\}$. Throughout,
parameters are positive and the interior equilibrium exists
($\Phi(O_0)>0$, Theorem~\ref{thm:uniqueness}).

\subsection{Step 1: the trapping rectangle, and why the old one leaked}

Let $\Ctrap$ be defined by Eq.~\eqref{eq:ctrap}, i.e.\ the unique capital
level at which the capital nullcline reaches the vitality ceiling,
$V_c(O(\Ctrap))=\Vmax$; at baseline
$\Ctrap=\bigl((0.1\cdot10/0.2-1)/0.15-1\bigr)/0.25=102.67$. Set
$\Omega=[0,\Vmax]\times[0,\Ctrap]$.

\emph{Forward invariance.} On the four faces of $\Omega$:
\begin{itemize}
\item $V=0$: $\dot V=R>0$ (inward; Appendix~\ref{app:invariance}).
\item $V=\Vmax$: $\dot V=-\delta O^{\gamma}\Vmax/(\Vmax+\varepsilon)<0$
(inward).
\item $C=0$: $\dot C=0$ (invariant face).
\item $C=\Ctrap$: for every $V\in[0,\Vmax]$,
\[
\dot C
= C\Bigl(\frac{\alpha V}{1+\varphi O(\Ctrap)}-\mu\Bigr)
\le C\Bigl(\frac{\alpha \Vmax}{1+\varphi O(\Ctrap)}-\mu\Bigr)
= 0,
\]
with equality only at the corner $V=\Vmax$, where the $V$-component points
strictly inward; no trajectory can exit through this face.
\end{itemize}
Moreover $\Omega$ is \emph{absorbing} for the region of interest: if
$C_0>\Ctrap$ (with $V_0\in[0,\Vmax]$), then $O>O(\Ctrap)$ implies
$\alpha V/(1+\varphi O)<\mu$, so $\dot C<0$ strictly until the trajectory
enters $\Omega$; hence every trajectory with $V_0\in[0,\Vmax]$, $C_0>0$
eventually lies in $\Omega$.

\emph{The correction.} Earlier drafts used the rectangle
$[0,\Vmax]\times[0,\Cmax]$ with the carrying capacity $\Cmax=44.99$ of
Eq.~\eqref{eq:cmax} as ceiling. That rectangle is \emph{not} forward
invariant: at $(V,C)=(\Vmax,44.99)$, $O=12.25$ and
\[
\dot C = C\Bigl(\frac{\alpha\Vmax}{1+\varphi O}-\mu\Bigr)
       = 44.99\,(0.3525-0.2) \approx +6.86 \;>\;0 ,
\]
so the flow exits through the ceiling; concretely, the baseline trajectory
from $(V_0,C_0)=(10,40)$ overshoots to $C\approx46.4$ before returning.
The carrying capacity is a flux threshold (Appendix~\ref{app:carrying});
the trapping bound is the nullcline-ceiling constant $\Ctrap$, and the two
must not be conflated. All results downstream of the trapping rectangle
are stated with $\Ctrap$; the certificate is verified numerically on
$[0,\Vmax]\times[0,\,1.2\,\Ctrap]$ by
\texttt{test\_bendixson\_dulac\_trapping\_rectangle\_is\_valid}.

\subsection{Step 2: Bendixson--Dulac certificate (no closed orbits)}

Take the Dulac function $B(V,C)=1/C$ on the open quadrant
$Q=\{V>0,\,C>0\}$, and write the field as
$\mathbf{F}=(F,G)$ with $G=C\,g(V,O)$, $g=\alpha V/(1+\varphi O)-\mu$.
Then
\[
\nabla\!\cdot\!\bigl(B\mathbf{F}\bigr)
= \frac{\partial}{\partial V}\Bigl(\frac{F}{C}\Bigr)
+ \frac{\partial}{\partial C}\bigl(g\bigr)
= \frac{1}{C}\Bigl(
   -\frac{R}{\Vmax}
   -\frac{\delta O^{\gamma}\varepsilon}{(V+\varepsilon)^{2}}
  \Bigr)
\;-\;
\frac{\alpha V\varphi\,\beta\eta C^{\eta-1}}{(1+\varphi O)^{2}} .
\]
The first bracket is strictly negative everywhere; the second term is
non-positive (strictly negative for $V>0$). Hence
$\nabla\!\cdot\!(B\mathbf{F})<0$ on all of $Q$---no sign changes, no
degenerate set. By the Bendixson--Dulac theorem
\citep{strogatz2015nonlinear}, $Q$ contains no closed orbit; the same
Green's-theorem argument excludes graphics (closed curves assembled from
orbits and equilibria) whose interior lies in $Q$. Numerically, the
supremum of the divergence over a dense grid in
$[0,\Vmax]\times(0,\,1.2\,\Ctrap]$ is $\approx-3.5\times10^{-3}$, bounded
away from zero
(\texttt{test\_bendixson\_dulac\_divergence\_strictly\_negative}).

\subsection{Step 3: equilibria inside $\Omega$}

By Theorem~\ref{thm:uniqueness} there is exactly one interior equilibrium
$(V^{*},C^{*})$; by the nullcline bound of Appendix~\ref{app:equilibrium}
it satisfies $C^{*}<\Ctrap$, so it lies in the interior of $\Omega$. By
Lemma~\ref{lem:trace} (trace negative, determinant positive) it is locally
asymptotically stable---at baseline a spiral sink
(Appendix~\ref{app:jacobian}). The only other equilibrium in $\Omega$ is
the axis point $(V_{\mathrm{axis}},0)$, $V_{\mathrm{axis}}\approx9.934$,
which is a hyperbolic saddle whose stable manifold is contained in the
invariant axis $\{C=0\}$ (Appendix~\ref{app:regularization}).

\subsection{Step 4: Poincar\'e--Bendixson assembly}

Let $(V_0,C_0)\in\Omega$ with $C_0>0$. By Step~1 the forward orbit remains
in the compact set $\Omega$, and by invariance of $\{C=0\}$ together with
uniqueness of solutions it remains in $Q$ for all finite time. Its
$\omega$-limit set $L$ is nonempty, compact, connected, and invariant. By
the Poincar\'e--Bendixson theorem, $L$ is an equilibrium, a closed orbit,
or a graphic. Step~2 excludes closed orbits and graphics in $Q$. Could $L$
meet the axis? The only invariant subsets of the axis face within $\Omega$
are the saddle and its axis-internal stable connections; if $L$ contained
the saddle, then (since $L$ is not the saddle alone unless the orbit
converges to it) $L$ would have to contain an orbit converging to the
saddle---an orbit in the saddle's stable manifold, which lies in
$\{C=0\}$ and is unreachable from $Q$. And convergence of the interior
orbit itself to the saddle is likewise impossible for the same reason.
Hence $L=\{(V^{*},C^{*})\}$: every trajectory with $C_0>0$ converges to
the interior equilibrium, which is locally stable by Step~3. This proves
global asymptotic stability on the open set $\{C>0\}\cap\Omega$, and, via
the absorbing property of Step~1, on
$\{(V,C):V\in[0,\Vmax],\,C>0\}$. \hfill$\square$

\subsection{Remarks}

\begin{enumerate}
\item \emph{The basin is open, and that is sharp.} $\{C=0\}$ is invariant:
a leader with exactly zero capital stays at zero capital, and the axis
flow converges to $(9.934,\,0)$---full vitality, no career. The theorem
therefore cannot be strengthened to the closed quadrant, and the
low-vitality equilibrium's basin is exactly $\{C>0\}$ within $\Omega$.
\item \emph{Why not a Lyapunov function.} No elementary global Lyapunov
function appears to exist: weighted quadratic forms
$a(V-V^{*})^{2}+b(C-C^{*})^{2}$ fail because the indefinite cross terms
generated by the $O^{\gamma}$ drain and the $1/(1+\varphi O)$ impact gate
cannot be dominated over all of $\Omega$ by any constant choice of
$(a,b)$. For a planar system, Dulac plus Poincar\'e--Bendixson is the
appropriate---not merely convenient---instrument; the certificate is also
mechanically checkable, which the companion tests exploit.
\item \emph{Numerical corroboration.} An $8\times8$ grid of initial
conditions spanning $[0.1,\Vmax]\times[0.5,\,2\Cmax]$ converges to
$(V^{*},C^{*})$ to within $10^{-2}$ in all 64 cases
(\texttt{test\_basin\_sweep\_all\_trajectories\_converge\_to\_zombie}),
providing an end-to-end check that no numerical artifact defeats the
analytical result.
\end{enumerate}
