% ============================================================================
% Chapter 5 — Discussion
% ============================================================================
\section{Discussion}
\label{sec:discussion}

\subsection{Theoretical contributions}
\label{sec:contributions}

DLVT contributes a closed-form, individual-level counterpart to the
organizational capability-trap literature. Where
\citet{repenning2002capability} and \citet{rahmandad2016capability} derive
bistable tipping structures---two basins, a separatrix, hysteresis---DLVT
shows that when the trap mechanism is routed through a single leader's
energetic budget, the structure collapses to one globally asymptotically
stable attractor: the trap is the default destination
(Theorem~\ref{thm:gas}). The practical difference is stark. In a bistable
world, prevention means staying in the healthy basin and rescue means a
large one-time push. In DLVT's world there is no healthy basin to defend;
only \emph{parameter} changes relocate the attractor, and
Corollary~\ref{cor:scope} says which parameters can do so. Relatedly, the
theory extends---rather than contests---human-capital theory
\citep{becker1993human} by adding the energetic-deployment constraint: the
wedge between capital owned and capital usable is the model's central
object, formalized in the energy-gated impact function \eqref{eq:impact}.

\subsection{Managerial implications}
\label{sec:implications}

Within the declared scope conditions, the model's implications invert a
common instinct. Scope reduction---delegating business units, flattening
portfolios, pruning interfaces, trimming spans of control---feels like the
natural remedy for executive overload. Within the power-law kernel family,
the model says it is absorbed: after a transient improvement, capital
re-expands until the complexity load is restored (P3), and equilibrium
vitality returns to where it was. The practical signature is the
delegation that relieves load for a quarter and then quietly refills. We
stress the hedge: this is a within-family prediction, and P4 marks where
it fails---where drain saturates in complexity, scope redesign regains
leverage.

What moves the equilibrium, in the model, is the recovery side: protected
restoration capacity ($R$), slower capital decay for the same impact
($\mu$), or higher conversion of impact into capital ($\alpha$).
Concretely, this ranks structural interventions---detachment blocks and
recovery routines built into the operating calendar, sabbatical
scheduling with protected re-entry---above one-time resets, consistent
with evidence that respite interventions produce real but decaying gains
unless structurally maintained \citep{davidson2010sabbatical} and with the
recovery literature's emphasis on ongoing rather than episodic restoration
\citep{sonnentag2007recovery,sonnentag2022recovery}. The prescription
converges with practitioner arguments for managing energy rather than
time \citep{loehr2003power,schwartz2007manage}, but it is derived here
from equilibrium structure rather than from anecdote, and it comes with a
falsifier attached (P6).

For boards and senior-talent processes, the model adds a caution about
inference from visible durability: at the low-vitality/high-status
equilibrium, status ($C^{*}$) is at its lifetime maximum precisely while
deployable energy is chronically below threshold, so an incumbent's
retained authority and reputation are uninformative about sustainability.
Pending the field test of \S\ref{sec:empirical}, all three implications
should be read as predictions to be audited, not prescriptions to be
rolled out.

\subsection{Limitations}
\label{sec:limitations}

Six boundaries delimit the theory's claims. Each is a deliberate scope
decision: we state what the model excludes, why, and what applies instead.

\paragraph{Deterministic dynamics.} The model is deterministic by design;
it has no stochastic forcing. Real careers experience shocks---health
events, market dislocations, reorganizations---that could produce
excursions and regime misclassification in finite samples. The
global-stability result guarantees return to the attractor but prices no
excursion costs; a stochastic extension is the natural next model, not a
patch on this one.

\paragraph{Single leader, closed loop.} Complexity is modeled as a
deterministic image of the leader's own capital
(Assumption~\ref{ass:scope}). This is a declared scope condition, not an
empirical claim: peer competition, organizational design, and market
turbulence all generate load that is orthogonal to own-capital
accumulation. Where exogenous load dominates, the exogenous-complexity
variant (Proposition~\ref{prop:band}) is the right object, and the
endogenous-feedback results do not apply.

\paragraph{Illustrative calibration.} Table~\ref{tab:parameters} has no
empirical anchor, and we quantify exactly how much hangs on it. The
baseline sits near its own regime boundary: $\mu/\alpha=2.0$ against a
flip value of $2.163$ ($8\%$ margin), and
$P(\text{low-vitality}\mid\text{stable equilibrium})\approx0.49$ across a
$\pm2\times$ log-uniform hypercube. Qualitative structure---uniqueness,
stability, scope absorption---is calibration-independent; regime labels
are not, and no baseline label in this paper should be read otherwise.

\paragraph{Stipulated threshold.} The strategic threshold
$\Vstrat=0.5\,\Vmax$ (Definition~\ref{def:threshold}) is a labeling
convention. The baseline equilibrium sits only $5.9\%$ below it, and any
threshold below $0.47\,\Vmax$ reclassifies the baseline as sustainable.
Classification claims are therefore conditional on $\theta$; empirical
work should estimate the threshold rather than inherit it.

\paragraph{Identifiability.} The eleven parameters collapse to
approximately six effective dimensionless groups after nondimensionalizing
$V$ by $\Vmax$, $C$ by $\beta^{-1/\eta}$, and time by $\mu^{-1}$; within
the equilibrium predictions there are pronounced ridges---most importantly
in $\mu/\alpha$ (which pins $V^{*}$ via \eqref{eq:vstar}) and $R/\delta$
(which pins the flux threshold)---so cross-sectional equilibrium data
alone cannot identify the full parameter vector. Identification requires
trajectories or shocks, and the empirical program is built accordingly
(\S\ref{sec:empirical}).

\paragraph{Level of formal guarantee.} The global-stability proof is
planar machinery (Dulac, Poincar\'e--Bendixson) and does not survive added
state dimensions. Extensions---teams, stochastic load, a second resource
stock---will need different tools, and we make no claims beyond two
dimensions.

\subsection{An empirical validation roadmap}
\label{sec:roadmap}

The propositions of \S\ref{sec:propositions} order the empirical program
from cheap to decisive. First, \emph{measurement and trajectory fitting}:
longitudinal administration of validated vitality and recovery instruments
\citep{ryan1997subjective,sonnentag2007recovery}, with career capital
proxied by scope, compensation, and network centrality, fitted in
continuous time to test the damped-spiral signature and single-attractor
convergence (P1). Second, \emph{exogenous shocks}: promotions,
reorganizations, and mandated leaves act as natural experiments that break
the $\mu/\alpha$ ridge which cross-sectional equilibrium data cannot (P2;
\S\ref{sec:limitations}). Third, the \emph{killer test}: a pre-registered
$2{\times}2$ field experiment crossing scope redesign with recovery
support over a 12--18~month horizon, long enough for the model's
re-expansion dynamics to complete. DLVT predicts a durable main effect of
recovery support on vitality, no durable main effect of scope redesign,
and conservation of the scope--capital product $\beta C^{\eta}$ in the
redesign arms (P3, P6); a durable vitality gain from scope redesign alone,
with the product not conserved, falsifies the central mechanism rather
than a calibration choice. Section~\ref{sec:empirical} develops this
program in full---measurement model, design cadence, estimation strategy,
identification plan, and the pre-registered protocol of the killer test.

\subsection{Conclusion}
\label{sec:conclusion}

Leadership research has documented, at length, that success is
energetically expensive \citep{hambrick2005executive,rudolph2020healthy}.
DLVT's contribution is to show that a minimal, fully analyzable model of
that observation---capital builds complexity, complexity drains vitality,
vitality gates impact---delivers strong and falsifiable structure: one
equilibrium, globally stable, low-vitality by default under convex load,
indifferent to scope redesign within its kernel family, and movable only
from the recovery side. The corrected mathematics in this manuscript is
deliberately less dramatic than earlier drafts: no finite-time collapse,
no critical coupling, no fold catastrophe. What remains is, we believe,
more useful: a small set of sharp claims, each with a stated boundary and
a stated way to kill it.
